### 2021-11-14 NSF2021/Description/specificaim1.tex
fixed highlighted sentence

The hydrophobic interactions come from the tendency of particles to
minimize the free energy of the structure of the surrounding water
molecules. The free energy functional takes the form 
\begin{align*}
\label{eq:free_energy}
  F[u] = C \int_{\Omega} \left(\rho |\nabla u|^2 + \rho^{-1} f(u)\right)
  \,d\xx,
\end{align*}
where $u$ is an order parameter for the structure of water, $\rho$ is a
decay length, $C$ is a constant, and $f(u)$ is a potential.
Hydrogen-bond persistence times are on the order of picoseconds
which is much smaller than characteristic time for 
particle motion \cite{MaGa13}.
Thus we assume $u$ minimizes $F[u]$ for all times.
Assuming appropriate conditions on $f(u)$ (\cite{evans10}, \S 8.2),
$u$ is bounded and satisfies the Euler-Lagrange equation
\begin{equation}
\label{eq:SL}
-\rho^2 \Delta u + \tfrac{1}{2}f'(u) = 0  \text{ in } \Omega,\quad u = g,
\text{ on } \bd\Omega.
\end{equation}
The boundary condition $g$ is a material label that is transported with
the particle (Figure~\ref{fig:bcs}).

@@

The formulation presents a number of mathematical and numerical
challenges. For one, the domain is constantly changing so that a new
boundary value problem must be solved at each time step. The particle
boundaries are nearly touching and high-order numerical
schemes and adaptively chosen time-step sizes
are required to calculate trajectories accurately.
Furthermore, the domain is unbounded and care must be taken 
in capturing the far-field boundary conditions correctly. 
Finally, the numerical routine must handle large system sizes
efficiently when the number of particles is large.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.9in}m{1.9in}m{1.9in}}            
      (a)
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.9in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.9in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{$t = t_{\text{final}}$}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:self-assembly}
    Rows (a)-(c) have the same initial configuration
    but have different boundary conditions $g.$ 
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results in a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.\\
  }
\end{figure}

\begin{wrapfigure}[14]{l}{2in}
  \vspace{-4pt}
  \centering
  \includegraphics[width=2in]{figures/SpecificAim1/OrderPDFs.jpg} 
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:OrderPDFs}
    Probability density functions for
    relative orientation of nearby particles
    for the data in 
    Figure~\ref{fig:self-assembly}, $t = t_{\text{final}}$.
  }
\end{wrapfigure}
Finally, boundary condition \eqref{eq:bcs}(c) models 
a \hl{particle with oppositelly charged ends} making the 
particle heads repel the tails \cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular to the
length of the chain. The equilibrium structure consists of particles
positioned on a square grid in a checkerboard pattern
Figure~\ref{fig:self-assembly}(c). This patterning is a consequence of
energy minimality and allows each particle to simultaneously coordinate
its head with the head of three other particles and its tail with the
tail of three other particles.

The ultimate goal is to use numerical simulations to aid
the syntheses of soft matter materials.
Here, mathematical models like the kind we propose are crucial
because the model parameters control surface properties independently, 
which can be difficult to accomplish experimentaly 
~\cite{Bradley2016,Mallory2017,Bradley2017}.
The ability to translate phenomenological parameters
into measurable quantities is important. 
For example, Figure~\ref{fig:OrderPDFs} plots
the probability density function for 
relative orientation of adjacent particles
for the three boundary condition cases:
(a) is unimodal with a peak at $\theta = 0$,
(b) is bimodal with peaks at $\theta = 0$ and $\theta = \pi$,
and (c) is multimodal with peaks at $\theta = 0$ and $\theta = \pm \pi/2$.
This kind of information can be used as an indicator
of the collective morphology. As another example,
it is costly or difficult to manufacture
Janus particles with uniform shape.
Thus, we can replace the idealistic boundary
conditions \eqref{eq:bcs} with a more realistic one like
\begin{equation}
\label{eq:vonMises}
  g_i(\xx) = \sqrt{
  \frac{\exp( \kappa_i\cos(\theta_i(\xx) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}.
\end{equation}
Here, the term under the square root is the von Mises distribution with
director angle $\theta^0_i$ and Bessel function $I_0$ of order zero~\cite{Fisher_1993}.
The square of this function is asymptotic to a normal distribution with variance
like $1/\kappa_i$, $\kappa_i \to \infty$. We can replicated surface imperfections
by randomizing the concentration $\kappa_i$ and particle radii $r_i$,
for example. \hl{Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions.}
In Specific Aim 3, we discuss how to include ions and surface charge.

@@

\subsubsection{Hydrophobic forces with double-well potential}
A natural generalization of the potential $f(u) = u^2$
is a fourth-order, asymmetric double-well potential
$f(u) = a(u-u_0)^2(u-u_1)^2 + c(u-u_0)^2 + bu$.
Here, the order parameter $u$ is related to
the mean number of nearest neighbors
per water molecule \cite{GoHaKo94}.
The value $u_0$ gives the absolute minimum in the bulk
and $u_1$ gives a local minimum at hydrophobic surfaces.
Figure \ref{fig:CA}(a) illustrates a double-well potential
and Specific Aim 2 describes the numerical implementation issues.

@@

Other approaches may be required if the modified phenomenological
theory does not recreate measurements.
One such approach is to consider Gaussian fluctuation. Here, we expand
the free energy
functional around the saddle point solution. The fluctuation energy
then involves calculation of the eigenfunctions of the second variation
of the free energy. This accounts for the fact that experimental
measurements are carried out at finite temperatures, where thermal
fluctuations should ideally also be considered. 
\todo[inline]{Above, needs citation and more solid explanation}

>>>>

The hydrophobic interactions come from the tendency of particles to
minimize the free energy of the structure of the surrounding water
molecules. The free energy functional takes the form 
\begin{equation}
\label{eq:free_energy}
  F[u] = C \int_{\Omega} \left(\rho |\nabla u|^2 + \rho^{-1} f(u)\right)
  \,d\xx,
\end{equation}
where $u$ is an order parameter for the structure of water, $\rho$ is a
decay length, $C$ is a constant, and $f(u)$ is a potential.
Hydrogen-bond persistence times are on the order of picoseconds
which is much smaller than characteristic time for 
particle motion \cite{MaGa13}.
Thus we assume $u$ minimizes $F[u]$ for all times.
Assuming appropriate conditions on $f(u)$ (\cite{evans10}, \S 8.2),
$u$ is bounded and satisfies the Euler-Lagrange equation
\begin{equation}
\label{eq:SL}
-\rho^2 \Delta u + \tfrac{1}{2}f'(u) = 0  \text{ in } \Omega,\quad u = g,
\text{ on } \bd\Omega.
\end{equation}
The boundary condition $g$ is a material label that is transported with
the particle (Figure~\ref{fig:bcs}).

@@

The formulation presents a number of mathematical and numerical
challenges. For one, the domain is constantly changing so that a new
boundary value problem must be solved at each time step. The particle
boundaries are nearly touching and high-order numerical
schemes and adaptively chosen time-step sizes
are required to calculate trajectories accurately.
Furthermore, the domain is unbounded and care must be taken 
in capturing the far-field boundary conditions correctly. 
Finally, the numerical routine must handle large system sizes
efficiently.

@@

\begin{wrapfigure}[16]{l}{2in}
  \vspace{-4pt}
  \centering
  \includegraphics[width=2in]{figures/SpecificAim1/OrderPDFs.jpg} 
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:OrderPDFs}
    Probability density functions for
    relative orientation of nearby particles
    for the data in 
    Figure~\ref{fig:self-assembly}, $t = t_{\text{final}}$.
  }
\end{wrapfigure}
Finally, boundary condition \eqref{eq:bcs}(c) models 
a particle whose head surface repels the tail surface as proposed in
\cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular to the
length of the chain. The equilibrium structure
resembles a checkerboard pattern
where each particle coordinates
its head with the head of three other particles and its tail with the
tail of three other particles Figure~\ref{fig:self-assembly}(c).

\begin{figure}[h!]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{$t = t_{\text{final}}$}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize
    \label{fig:self-assembly}
    Rows (a)-(c) have the same initial configuration
    but have different boundary conditions $g.$ 
    In row (a), self-assembly with an amphiphilic boundary
    condition \eqref{eq:bcs}(a) results in a bilayer pattern shielding
    the hydrophobic core (yellow, $u > 0$) from the aqueous phase (blue, $u = 0$).
    Row (b) shows particles with a hydrophobic intensity that is
    greater on one side of the particle (magenta) than the other (yellow).
    In row (c), green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    Particles orient like sides to form a checkerboard pattern.\\
  }
\end{figure}

The ultimate goal is to use numerical simulations to aid
the syntheses of soft matter materials.
Here, mathematical models like the kind we propose are crucial
because the model parameters control surface properties independently, 
which can be difficult to accomplish experimentaly 
~\cite{Bradley2016,Mallory2017,Bradley2017}.
The ability to translate phenomenological parameters
into measurable quantities is important. 
For example, Figure~\ref{fig:OrderPDFs} plots
the probability density function for 
relative orientation of adjacent particles
for the three boundary condition cases:
(a) is unimodal with a peak at $\theta = 0$,
(b) is bimodal with peaks at $\theta = 0$ and $\theta = \pi$,
and (c) is multimodal with peaks at $\theta = 0$ and $\theta = \pm \pi/2$.
This kind of information can be used as an indicator
of the collective morphology. As another example,
it is costly or difficult to manufacture
Janus particles with uniform shape.
Thus, we can replace the idealistic boundary
conditions \eqref{eq:bcs} with a more realistic one like
\begin{equation}
\label{eq:vonMises}
  g_i(\xx) = \sqrt{
  \frac{\exp( \kappa_i\cos(\theta_i(\xx) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}.
\end{equation}
Here, the term under the square root is the von Mises distribution with
director angle $\theta^0_i$ and Bessel function $I_0$ of order zero~\cite{Fisher_1993}.
The square of this function is asymptotic to a normal distribution with variance
like $1/\kappa_i$, $\kappa_i \to \infty$. We can replicated surface imperfections
by randomizing the concentration $\kappa_i$ and particle radii $r_i$,
for example.
%\hl{Collaboration with experimental groups will
%use numerical methods to predict parameter regimes 
%to control particle-particle interactions.}

@@

\subsubsection{Hydrophobic forces with double-well potential}
A natural generalization of the potential $f(u) = u^2$
is a fourth-order, asymmetric double-well potential
$f(u) = a(u-u_0)^2(u-u_1)^2 + c(u-u_0)^2 + bu$.
Here, the order parameter $u$ is related to
the mean number of nearest neighbors
per water molecule \cite{GoHaKo94}.
The value $u_0$ gives the absolute minimum in the bulk
and $u_1$ gives a local minimum at hydrophobic surfaces.
Figure \ref{fig:CA}(a) illustrates a double-well potential.

@@

Specific Aim 2 describes how we plan to numerically deal with nonlinear
equations. Once we have established implementations, 
we can consider further generalizations of 
\eqref{eq:free_energy} to the Oseen-Frank energy for a nematic liquid crystal
~\cite{doi:10.1137/15M103844X, Bartels,AIHPA_1997__66_4_411_0,
 deGennesProst,HaKiLi86,Virga}.  
For now, Specific Aim 3 describes in detail how we plan to handle
charges and ions in solution.

%Other approaches may be required if the modified phenomenological
%theory does not recreate measurements.
%One such approach is to consider Gaussian fluctuation. Here, we expand
%the free energy
%functional around the saddle point solution. The fluctuation energy
%then involves calculation of the eigenfunctions of the second variation
%of the free energy. This accounts for the fact that experimental
%measurements are carried out at finite temperatures, where thermal
%fluctuations should ideally also be considered. 
%\todo[inline]{Above, needs citation and more solid explanation}
